<html>
<head>
<link rel="stylesheet" type="text/css" href="style.css">
</head>

<BODY text=#000000 bgColor=#ffffff leftMargin=0 topMargin=0>
<TABLE height="100%" cellSpacing=0 cellPadding=5 width="100%" border=0>
 
  <TR>
      <TD class=sidebar align=middle valign=top><h3>
<a href="index.html">Meeting Home</a><p>
<a href="sciprog.html">Science Program</a><p>
<a href="venue.html">Venue</a><p>
<a href="travel.html">Transportation</a><p>
<a href="acc.html">Accommodations</a><p>
<a href="abs.html">Abstract Submission</a><p>
<a href="people.html">List of Participants</a><p>
<a href="local.html">Local Guides</a><p>
</h3>
      </td>
      <td>
<table>
<td><h2> #xB5AB </h2> <br>
<br>
<hr>
<br>
Workshop oral sessions are expected to run from the morning of July 27 through the afternoon of July 29, so we recommend arriving at Ithaca on July 26 and departing
on July 30.
<br>
<br>
<h3>Airline Travel</h3>
<P>The closest airport to the meeting site is the <a href="http://www.flyithaca.com/">Ithaca Tompkins Regional Airport</a> (airline code ITH), approximately a 10-minute drive from campus and downtown.  This is a small regional airport (and thus much more friendly and convenient than larger airports) with only three airlines flying into it:  US Airways (from Philadelphia and New York LaGuardia), Delta (from Detroit), and Continental (from Newark).  Good fares into ITH can often be found, especially if you reserve in advance, and the airline coverage includes two of the three major airline networks (Star Alliance and SkyTeam).

<p>The closest mid-size airports, each with much broader airline coverage, are <a href="http://www.syrairport.org/">Syracuse Hancock International Airport</a> (SYR) approximately 60 miles north of Ithaca with an estimated driving time of 1-1/4 hours, and <a href="http://www.monroecounty.gov/airport-index.php">Greater Rochester International Airport</a> (ROC) approximately 100 miles northwest of Ithaca with an estimated driving time of just over 2 hours.  International travelers flying into New York City may elect to rent a car; the drive is 4+ hours, not including NYC traffic.

<p>Recommended airport transportation options vary from airport to airport.  

<ul>
<li><b>Ithaca (ITH):</b> <a href="http://www.ithacataxi.biz/">Ithaca Dispatch</a> charges $16 to $18 for the first passenger, and $4/person (if family) or $8/person (if not family) thereafter.  Check whether your hotel provides a free shuttle.  
<li><b>Syracuse (SYR):</b> <a href="http://www.ithacaairlinelimo.org/">Ithaca Airline Limousine</a> charges $80/person one-way or $110/person round trip. 
<li><b>Rochester (ROC):</b> <a href="http://www.ithacataxi.biz/">Ithaca Dispatch</a> charges $225 for the first two passengers, $20/person thereafter.
</ul>

<h3>Getting Around in Ithaca</h3>

<p>Most of the major hotels have their own shuttle buses that are available for local travel.  Also, <a href="http://www.tcatbus.com">Tompkins Consolidated Area Transit</a> (TCAT) is an excellent public bus system.

<h3>Parking</h3>

<p>Parking on the Cornell campus is regulated Monday through Friday, 7:30am to 5:00pm. A guest parking pass can be purchased at the ticket booths, located at every major campus entrance, for $10 (one day), or $6 (four hours).  To avoid parking charges, transportation to Cornell from most hotels is available either by hotel shuttle or by TCAT bus.  More information to appear later. 

<!--<h3>Driving to and in Ithaca</h3>

<p>Ithaca is divided into topographic regions by several deep valleys,
and driving between ostensibly nearby points can be tricky due to steep
grades traversed only by minor roads. Often the best driving route is
to go into Downtown and then out again. It is generally a good idea to
know where in Ithaca you are going, and plan your route beforehand.
Ithaca's major regions are as follows:
</p><ul><li>The main valley that includes Cayuga Lake and Downtown continues to the southwest along Route 13 South (<b>Elmira Road</b>, including hotels #3, #7, and #13).  
</li><li>Nearby but on much higher ground is <b>South Hill</b>, served by Route 96B (Danby Road, including hotel #4).  Ithaca College is here. 
</li><li>Cornell is on <b>East Hill</b>, served by Routes 79 and 366 (including hotel #1).

</li><li>The airport and a large commercial district are north of Ithaca, near Route 13 North and <b>Triphammer Road</b> (including hotels #2, #5, #6, #10, and #11).  
</li><li>West of the main valley is <b>West Hill</b>, served by Routes 79, 96, and 89, and including the Hospital and the Museum of the Earth (banquet site). 
</li></ul>
<p>Many people <b>driving to Ithaca from points south and east</b> will
pass through the convergence of I-81, I-86, and I-88 near Binghamton,
NY. The best route from Binghamton to Ithaca depends on where in Ithaca
you are going; here are some Google Maps that have best routes
specified: </p><ul><li><a href="http://maps.google.com/maps?f=d&amp;hl=en&amp;geocode=13683620223581194179,42.115520,-75.905360%3B16105807137725549934,42.364250,-76.113040%3B14413553406352876767,42.413620,-76.372590&amp;saddr=I-81+N%2FRT-17+W+%4042.115520,+-75.905360&amp;daddr=RT-79+%4042.364250,+-76.113040+to:42.39608,-76.392059+to:Ithaca,+NY&amp;mra=dme&amp;mrcr=0&amp;mrsp=2&amp;sz=10&amp;via=1,2&amp;sll=42.278005,-76.19549&amp;sspn=0.430817,0.85144&amp;ie=UTF8&amp;z=10">to East Hill and Downtown</a> (I-81 to NY-79 West)
</li><li><a href="http://maps.google.com/maps?f=d&amp;hl=en&amp;geocode=13683620223581194179,42.115520,-75.905360%3B17170984054280147295,42.360252,-76.091516%3B16045593399591524202,42.445990,-76.184890&amp;saddr=I-81+N%2FRT-17+W+%4042.115520,+-75.905360&amp;daddr=RT-79+%4042.360252,+-76.091516+to:42.426498,-76.231384+to:North+Triphammer+Road,+Ithaca,+NY&amp;mra=dme&amp;mrcr=0&amp;mrsp=2&amp;sz=10&amp;via=1,2&amp;sll=42.306975,-76.19297&amp;sspn=0.430623,0.85144&amp;ie=UTF8&amp;z=10">to Triphammer area</a> (I-81 to NY-79 West to NY-38 North to NY-13 South/West)

</li><li><a href="http://maps.google.com/maps?f=d&amp;hl=en&amp;geocode=13683620223581194179,42.115520,-75.905360%3B16954071926535123347,42.222550,-76.405400&amp;saddr=I-81+N%2FRT-17+W+%4042.115520,+-75.905360&amp;daddr=42.278325,-76.374207+to:Danby+Road,+Ithaca,+NY&amp;mra=dme&amp;mrcr=0&amp;mrsp=1&amp;sz=10&amp;via=1&amp;sll=42.218245,-76.197675&amp;sspn=0.431227,0.85144&amp;ie=UTF8&amp;z=10">to South Hill</a> (I-86/NY-17 West to NY-96 North to NY-96B North)
</li><li><a href="http://maps.google.com/maps?f=d&saddr=I-81+N%2FRT-17+W+%4042.115520,+-75.905360&daddr=42.212245,-76.448364+to:Elmira+Rd,+Ithaca,+NY+14850&hl=en&geocode=FcChggIdsMZ5-w%3B%3B&mra=dme&mrcr=0&mrsp=1&sz=10&via=1&sll=42.237669,-76.230011&sspn=0.618174,1.29776&ie=UTF8&t=h&z=10">to Elmira Road</a> (I-86/NY-17 West to NY-96 North to NY-13 North). 
</li></ul>
<p>An <b>alternate route to Ithaca from points south</b> is to leave
the Interstate at Harrisburg, PA, and take US-15 North through Corning,
NY. This route is shorter by miles and only slightly longer by time,
and can be less stressful and more scenic. <a href="http://maps.google.com/maps?f=d&amp;hl=en&amp;geocode=1918333228698884833,40.982740,-76.882180&amp;saddr=Harrisburg,+PA&amp;daddr=W+Branch+Hwy%2FUS-15+N+%4040.982740,+-76.882180+to:41.157978,-76.799927+to:Ithaca,+NY&amp;mra=dpe&amp;mrcr=0&amp;mrsp=2&amp;sz=9&amp;via=1,2&amp;sll=40.890677,-76.398926&amp;sspn=0.880352,1.702881&amp;ie=UTF8&amp;ll=41.58258,-76.794434&amp;spn=3.484221,6.811523&amp;z=7">Here is a map</a> of that route.  
-->

<br>
<br>
<br>
<br>

</td>
</tr>

</body>
